Making (Secure) FETCH Happen: \\ \vspace{2 mm} {\large End-to-End Metadata Security for IMAP}

### 2015-04-29 paper/paper.tex
various updates to the paper

\section{Design overview}
To meet all aforementioned requirements, we present \projectnospace, a client-only extension to the IMAP protocol that provides metadata privacy-preserving storage for multiple \project clients. In \projectnospace, emails received at the client are fully encrypted (including mail headers), indexed, and uploaded to a specialized mailbox in the IMAP server. This hides sender and recipient information from the IMAP server. Encryption is based on a passphrase and generated salt, using scrypt as a key derivation function. A special-use ``INDEX'' message is stored in the mailbox as well, and tracks metadata for the client such as the apparent mailbox that a message is in (called a pseudo-mailbox, and visible only to clients, not the server). A special-use ``LOCK'' message is used to allow multiple clients access to the same mailbox, in keeping with IMAP's design principles. Message arrival times are hidden from the server using a combination of scheduling, mixing, fake decoy messages, and techniques drawn from oblivious RAM. \project provides rollback detection for both messages and the index, using version numbers saved client-side and message hashes stored server-side. It is robust to client failure in any state.

@@

@@

When a message is received at the IMAP Server as in Section~\ref{mail-from-insecure}, \project processes the message so it bears surface equivalence to messages that arrived at the client machine from a secure source. Note that full equivalence can only be achieved by also including mixing steps either here or during scheduled message exchanges. Without them, it is trivial for an IMAP server to keep track of which message in CRYPTOBLOBS corresponds to which message that was received in the INBOX, because the server will see that a message was added to CRYPTOBLOBS and deleted from INBOX at about the same time.

@@

\subsection{Deletion}
Various \project actions involve deletion of a message from the CRYPTOBLOBS mailbox. In general, not deleting a message will not leak any data to the IMAP server, since our threat model includes a server than can either refuse to delete or can save a backup version of any message. Yet for practical reasons such as server data usage and lack of clutter, it is generally beneficial to at least allow a client to attempt to delete a message. Here we must note that IMAP servers may differ in their implementation in regard to the location of messages, potentially complicating deletion protocols. For example, Gmail uses its All Mail mailbox (``label'', in their terms) as a message archive \cite{allmail}. That is, if a message appears in any mailbox, it will also appear in the All Mail mailbox. If it is deleted from All Mail but still exists elsewhere, it will be replaced in All Mail. Thus, to delete a message from any non-All Mail mailbox, it must first be deleted from all other mailboxes then deleted from All Mail. Note that the message will be assigned a different UID for each mailbox that it appears in.

@@

When a message is appended to the CRYPTOBLOBS mailbox, whether it be a processed INBOX message, a fake message created on schedule, or a message from a metadata-secure source, it is encrypted before being uploaded. It is encrypted using the standard method of authenticated encryption, which is to say that a Message Authentication Code computed on the encrypted message is concatenated to the encrypted message. Data is encrypted using AES-CBC and signed with HMAC-SHA256. The encryption key is derived from the random salt created during initialization (as described in Section~\ref{initialization}) and a passphrase using scrypt \cite{percival2009stronger}. Scrypt is a memory-hard key derivation function that imposes a cost for deriving a key given a guess of a passphrase by using many sequential iterations that cannot be parallelized on specialized hardware.

@@

\section{Protecting Metadata}
Once a message has been received at the client, we would like to upload it to the IMAP server for storage. Yet while we can hide the FROM and TO headers easily, we do not want the server to gain timing information about the message. If we upload it immediately, then the server will know that we have received a message at that time. Even worse, modern clients often save sent messages to a Sent Messages mailbox. If the mail server of the sender and the receiver are colluding, as can easily be the case when both clients are using the same company to provide servers, then the server could easily execute timing attacks by observing two users suspected to be communicating, and seeing if they upload a message at about the same time. So, we must include a mechanism for hiding this timing metadata from the IMAP server. The solution must not:

@@

\subsection{Mechanisms}
While we can vary particular parameters and algorithms, most potential solutions discussed in this section rely on two primary components, scheduling and mixing. By scheduling, we mean that the client will exchange data with the server on a set or randomized schedule. By mixing, we mean that in an exchange with the server, multiple messages will be deleted, and multiple messages uploaded in their place, with only the client being aware of the correlations between messages deleted and uploaded.

>>>>

\section{Design overview}
To meet all aforementioned requirements, we present \projectnospace, a client-only extension to the IMAP protocol that provides metadata privacy-preserving storage for multiple \project clients. In \projectnospace, emails received at the client are fully encrypted (including mail headers), indexed, and uploaded to a specialized mailbox in the IMAP server. This hides sender and recipient information from the IMAP server. Encryption is based on a passphrase and generated salt, using scrypt as a key derivation function. A special-use ``INDEX'' message is stored in the mailbox as well, and tracks metadata for the client such as the apparent mailbox that a message is in (called a pseudo-mailbox, and visible only to clients, not the server). A special-use ``LOCK'' message is used to allow multiple clients access to the same mailbox, in keeping with IMAP's design principles. Message arrival times are hidden from the server using a combination of scheduling, mixing, fake decoy messages, and techniques drawn from oblivious RAM. \project provides rollback detection for both messages and the index, using version numbers saved client-side and message hashes stored server-side. It is robust to client failure in any state. Details of the design will appear in later sections.

@@

All modified commands rely on the metadata INDEX, the store for all cross-client data. The INDEX is a dictionary that is serialized as a string, encrypted, and saved as a message on the IMAP server. All \project clients can access the INDEX; it is their primary medium for synchronizing changes. The INDEX contains a wealth of information. First, its subject line includes the salt associated with the user's account that all clients will require (in conjunction with a passphrase) to decrypt any data stored in the server. Second, it contains a SEQUENCE NUMBER used to detect rollback. Third, for each mailbox known to the client or only internally to the \project module, it contains a dictionary mapping the message ID as it is appears to the client to the UNIQUE SUBJECT number used by \project to locate the message, along with a hash of the message contents. Fourth, for each folder known to the client but not to the server, it contains a NEXT UID field that stores the next client-facing message ID to be assigned.

@@

When a message is received at the IMAP Server as in Section~\ref{mail-from-insecure}, \project processes the message so it bears surface equivalence to messages that arrived at the client machine from a secure source. Note that full equivalence can only be achieved by also including mixing steps either here or during scheduled message exchanges. Without them, it is trivial for an IMAP server to keep track of which message in CRYPTOBLOBS corresponds to which message that was received in the INBOX, because the server will see that a message was added to CRYPTOBLOBS and deleted from INBOX at about the same time. Mixing will be further explored in Section~\ref{metadatasecurity}.

@@

\subsection{Deletion}
Various \project actions involve deletion of a message from the CRYPTOBLOBS mailbox. In general, although a user might want to delete a message for practical reasons such as server data usage and lack of clutter, deleting a message from the server provides no additional theoretical security value, since our threat model includes a server than can either refuse to delete the message or can save its own hidden copy of it. That being said, we should still support a deletion command to ensure a continuous user experience, and for use in \project abstractions such as mixing. Here we must note that IMAP servers may differ in their implementation regarding the location of messages, potentially complicating deletion protocols. For example, Gmail uses its All Mail mailbox (``label'', in their terms) as a message archive \cite{allmail}. That is, if a message appears in any mailbox, it will also appear in the All Mail mailbox. If it is deleted from All Mail but still exists elsewhere, it will be replaced in All Mail. Thus, to delete a message from any non-All Mail mailbox, it must first be deleted from all other mailboxes then deleted from All Mail. Note that the message will be assigned a different UID for each mailbox that it appears in.

@@

When a message is appended to the CRYPTOBLOBS mailbox, whether it be a processed INBOX message, a fake message created on schedule, or a message from a metadata-secure source, it is encrypted before being uploaded. It is encrypted using the standard method of authenticated encryption, which is to say that a Message Authentication Code computed on the encrypted message is concatenated to the encrypted message. Data is encrypted using AES-CBC and signed with HMAC-SHA256. The encryption key is derived from the random salt created during initialization (as described in Section~\ref{initialization}) and a passphrase using scrypt \cite{percival2009stronger}. Scrypt is a memory-hard key derivation function that imposes a cost for deriving a key given a guess of a passphrase by using many sequential iterations that cannot be parallelized on specialized hardware. To support mixing operations, we must also store a nonce in the INDEX, and at each mixing step update the nonce and re-encrypt the message.

@@

\section{Protecting Metadata}
\label{metadatasecurity}

Once a message has been received at the client, we would like to upload it to the IMAP server for storage. Yet while we can hide the FROM and TO headers easily, we do not want the server to gain timing information about the message. If we upload it immediately, then the server will know that we have received a message at that time. Even worse, modern clients often save sent messages to a Sent Messages mailbox. If the mail server of the sender and the receiver are colluding, as can easily be the case when both clients are using the same company to provide servers, then the server could easily execute timing attacks by observing two users suspected to be communicating, and seeing if they upload a message at about the same time. So, we must include a mechanism for hiding this timing metadata from the IMAP server. The solution must not:

@@

\subsection{Mechanisms}

While we can vary particular parameters and algorithms, most potential solutions discussed in this section rely on two primary components, scheduling and mixing. By scheduling, we mean that the client will exchange data with the server on a set or randomized schedule. By mixing, we mean that in an exchange with the server, multiple messages will be deleted, and multiple messages uploaded in their place, with only the client being aware of the correlations between messages deleted and uploaded.
